Proving linearizability using forward simulations

### 2017-05-04 draftCAV/draft.tex
the abstract queue
changes on the abstract queue implementation
removing appendices
stack implementation

\lstset{
  language={Java},
  basicstyle=\ttfamily\scriptsize
}

@@

\title{Proving linearizability using forward simulations}
\author{ }
\institute{ } 
\begin{document}

@@

\vspace{-9mm}
\begin{abstract}
%Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
Linearizability is the standard correctness criterion concurrent data structures such as stacks and queues. %
%in this context. 
It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.
%or equivalently observational refinement, 
Proving linearizability requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward \emph{and} backward simulations. In both cases, carrying out proofs is hard and complex in general. In particular, backward reasoning is difficult in the context of programs with data structures, and strategies for identifying statically linearization points cannot be defined for all existing implementations.  In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy\&Wing Queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments. This leads to simple and natural correctness proofs for these implementations that are amenable to automation.

%Efficient implementations of atomic objects such as concurrent stacks and queues are especially susceptible to programming errors, and necessitate automatic verification. Unfortunately their correctness criteria -- conformance to atomic reference implementations -- are hard to verify. Conformance can be formalized as a trace inclusion problem, also known as observational refinement, which in general, can be proved using compositions of forward and backward simulations. Contradicting common belief, we show that many complex implementations, e.g., the Herlihy and Wing queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments. We define classes of queue and stack implementations for which forward simulations are sound and complete for proving their correctness.
\vspace{-2mm}
\end{abstract}

@@

@@

\newpage
\appendix

\input{app-prelim}
\input{app-backSim}
\input{app-absImplQueue}
\input{app-absImplStack}
\input{app-tss}

>>>>

\lstset{
  language={program},
  basicstyle=\ttfamily\scriptsize
}

@@

\title{Proving linearizability using forward simulations}
\author{Ahmed Bouajjani\inst{1}\and Michael Emmi\inst{2}\and Constantin Enea\inst{1} \and Suha Orhun Mutluergil\inst{3}}
\institute{IRIF, Univ. Paris Diderot, \email{\{abou,cenea\}@irif.fr} \and Nokia Bell Labs, \email{michael.emmi@nokia.com}\and Koc University, \email{smutluergil@ku.edu.tr}}
\begin{document}

@@

%\vspace{-9mm}
\begin{abstract}
%Efficient concurrent implementations of abstract data structures such as stacks and queues are especially susceptible to programming errors, and necessitate automatic verification.
Linearizability is the standard correctness criterion for concurrent data structures such as stacks and queues. %
%in this context. 
It allows to establish observational refinement between a concurrent implementation and an atomic reference implementation.
%or equivalently observational refinement, 
Proving linearizability requires identifying linearization points for each method invocation along all possible computations, leading to valid sequential executions, or alternatively, establishing forward \emph{and} backward simulations. In both cases, carrying out proofs is hard and complex in general. In particular, backward reasoning is difficult in the context of programs with data structures, and strategies for identifying statically linearization points cannot be defined for all existing implementations.  In this paper, we show that, contrary to common belief, many such complex implementations, including, e.g., the Herlihy\&Wing Queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments. This leads to simple and natural correctness proofs for these implementations that are amenable to automation.

%Efficient implementations of atomic objects such as concurrent stacks and queues are especially susceptible to programming errors, and necessitate automatic verification. Unfortunately their correctness criteria -- conformance to atomic reference implementations -- are hard to verify. Conformance can be formalized as a trace inclusion problem, also known as observational refinement, which in general, can be proved using compositions of forward and backward simulations. Contradicting common belief, we show that many complex implementations, e.g., the Herlihy and Wing queue and the Time-Stamped Stack, can be proved correct using only forward simulation arguments. We define classes of queue and stack implementations for which forward simulations are sound and complete for proving their correctness.
%\vspace{-2mm}
\end{abstract}

@@

%\input{abstract_queues}

@@

%\newpage
%\appendix
%
%\input{app-prelim}
%\input{app-backSim}
%\input{app-absImplQueue}
%\input{app-absImplStack}
%\input{app-tss}

### 2017-05-04 draftCAV/queue.tex
changes on the abstract queue implementation
stack implementation

Since the linearization points of the enqueues are not fixed, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$. 
In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.

@@

\begin{wrapfigure}{l}{6.7cm}
\vspace{-6mm}
\includegraphics[width=6.8cm]{fig-queue12.pdf}
%
%\vspace{2mm}
%\includegraphics[width=7cm]{fig-queue2.pdf}
\vspace{-8mm}
\caption{Simulating queue histories with $AbsQ$. The order between actions is from left to right.}
\label{fig:queueSim}
\vspace{-6mm}
\end{wrapfigure}
Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points). The state of $AbsQ$ after each action is pictured as a graph below the action. The nodes of this graph represent enqueue operations and the edges happens-before constraints. Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed. For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$. Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.

For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}. The states of $AbsQ$ consist of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}. The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$. $Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
%
The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}, provided that the operation is still present in the current state. The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point. When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic. It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k} which is minimal in the current happens-before. In the latter case, $\mathsf{present}${\tt (k)} is set to {\tt false} to indicate that {\tt k}'s input has been removed. When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t. the current happens-before).

@@

\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
rule inv(enq,x,k) = {
  $\mathsf{present}$(k) = true;
  $\mathsf{pending}$(k) = true;
  $\mathsf{arg}$(k) = x;
  forall k' with $\mathsf{present}$(k)$\land \neg \mathsf{pending}$(k') do
    $\mathsf{before}$(k',k) = true;
}

rule lin(deq,y,k) = {
  choose a $\in$ {true, false};
  if ((a || $\forall k.\ \mathsf{present}(k)$=false) && $\forall k.\ \mathsf{present}(k)\Rightarrow \mathsf{pending}(k)$) 
    y = EMPTY;
  else 
    choose k $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k$)$;
    $\mathsf{present}$(k) = false;
    y = $\mathsf{arg}$(k);
}
\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-1mm}
\begin{lstlisting}
rule ret(enq,k) = {
  $\mathsf{pending}$(k) = false;
  return;
}
rule inv(deq,k) = {
  skip;
}
rule ret(deq,y,k) = {
  return y;
}
\end{lstlisting}
\end{minipage}
%	\caption{The abstract queue implementation $AbsQ$.}
%	\label{fig:signatures}
%\end{figure}

%{\scriptsize
%  \centering
%  \begin{mathpar}
%    \inferrule[call-enq]{
%      k\not\in dom(cp) \\ 
%      d\neq {\tt EMPTY}
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{inv(enq,d,k)}
%      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
%    }\hspace{5mm}
%
%    \inferrule[call-deq]{
%      k\not\in dom(cp) \\ 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{inv(deq,k)}
%      O,<,\ell,rv,cp[k\mapsto R_1]
%    }\hspace{5mm}
%    \inferrule[ret-deq]{
%       cp(k) = R_2 \\
%       rv(k)=d  
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(deq,d,k)}
%      O,<,\ell,rv,cp[k\mapsto R_3]
%    }\hspace{5mm}
%
%    \inferrule[ret-enq1]{
%      cp(k) = A_1 \\
%      k \in O \\
%      \ell(k) = (d,{\tt PEND}) 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(enq,k)}
%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
%    }\hspace{5mm}
%    \inferrule[ret-enq2]{
%      cp(k) = A_1 \\
%      k \not\in O 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(enq,k)}
%      O,<,\ell,rv,cp[k\mapsto A_2]
%    }\hspace{5mm}
%
%    \inferrule[lin-deq1]{
%       cp(k) = R_1 \\
%       d\neq{\tt EMPTY} \\
%       k'\in min(O) \\ 
%       \ell_1(k')=d
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{lin(deq,d,k)}
%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
%    }\hspace{5mm}
%    \inferrule[lin-deq2]{
%       cp(k) = R_1 \\
%       \forall o\in O.\ \ell_2(o)={\tt PEND}
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{lin(deq,{\tt EMPTY},k)}
%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
%    }\hspace{5mm}    
%      \end{mathpar}
%  }
 \vspace{-6mm}
  \caption{The abstract queue implementations $AbsQ$. Each macro rule is executed atomically in one step (indicated by the keyword {\ttfamily \bfseries atomic}).
%  transition relation of $AbsQ$. We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.
  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
  }
  \label{fig:transitions:AbsQ}
\vspace{-6mm}
\end{figure}

>>>>

Since the linearization points of the enqueues are determined by future dequeue invocations, there exists no forward simulation from $\mathit{HWQ}$ to $AbsQ_0$.  
In the following, we describe the abstract implementation $AbsQ$ for which such a forward simulation does exist.

@@

\begin{wrapfigure}{l}{6.7cm}
\vspace{-6mm}
\includegraphics[width=6.8cm]{fig-queue12.pdf}
%
%\vspace{2mm}
%\includegraphics[width=7cm]{fig-queue2.pdf}
\vspace{-8mm}
\caption{Simulating queue histories with $AbsQ$. An operation is pictured by a line delimited by two circles denoting the call and respectively, the return action. The representation of a dequeue operation includes a red circle that stands for a {\tt swap} returning a non-null value, which is their linearization point.}
\label{fig:queueSim}
\vspace{-6mm}
\end{wrapfigure}
Figure~\ref{fig:queueSim} pictures two executions of $AbsQ$ for two extended histories (that include dequeue linearization points). The state of $AbsQ$ after each action is pictured as a graph below the action. The nodes of this graph represent enqueue operations and the edges happens-before constraints. Each node is labeled by a value (the input of the enqueue) and a flag {\tt PEND} or {\tt COMP} showing whether the operation is pending or completed. For instance, in the case of the first history, the dequeue linearization point $lin(deq,y,3)$ is enabled because the current happens-before contains a \emph{minimal} enqueue operation with input $y$. Note that a linearization point $lin(deq,x,3)$ is also enabled at this state.

For readability, we define $AbsQ$ as an abstract state machine, which is given in Fig.~\ref{fig:transitions:AbsQ}. The states of $AbsQ$ consist of several boolean predicates indicating whether the value added by an enqueue has not been removed yet (the predicate $\mathsf{present})$, whether an enqueue is pending (the predicate $\mathsf{pending}$), the happens-before order (the predicate $\mathsf{before}$), and a function giving the value added by an enqueue (the function $\mathsf{arg}$).
%the states of $AbsQ$ are tuples $\tup{O,<,\ell,rv,cp}$ where $O\subseteq \<Ops>$ is a set of operation identifiers, $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a pending/completed flag (the flag is used to track the happens-before order), $rv:\<Ops> ~> \<Vals>$ records the return value of a dequeue fixed at its linearization point ($~>$ denotes a partial function), and $cp:\<Ops> ~> \{A_1,A_2,R_1,R_2,R_3\}$ records the control point of every enqueue ($A_1, A_2$) or dequeue operation ($R_1,R_2,R_3$).
%All the components are $\emptyset$ in the initial state, and the transition relation $->$ is defined in Fig.~\ref{fig:transitions:AbsQ}. The alphabet of $AbsQ$ contains call/return actions and dequeue linearization points, denoted by $lin(deq,d,k)$. $Lin(deq)$ is the set of all actions $lin(deq,d,k)$.
%
The enqueue operations consist of two macro rules: the rule {\tt inv(enq,x,k)} orders the invoked operation after all the completed enqueues present in the current state ({\tt k} is a fresh operation identifier associated to the current operation), and the rule {\tt ret(enq,k)} sets $\mathsf{pending}${\tt (k)} to {\tt false}. % provided that the operation is still present in the current state. 
The dequeue operations consist of two rules {\tt inv(deq,k)} and {\tt ret(deq,y,k)} associated to the invocation and respectively, the return action, whose definition is obvious, and the rule {\tt lin(deq,y,k)} corresponding to its linearization point. When a non-empty set of pending enqueues are present, the rule {\tt lin(deq,y,k)} is non-deterministic. It can set the return value {\tt y} either to {\tt EMPTY}, or to a value {\tt d} $\neq$ {\tt EMPTY}, provided that {\tt d} has been added by an enqueue {\tt k1} which is minimal in the current happens-before. In the latter case, $\mathsf{present}${\tt (k1)} is set to {\tt false} to indicate that {\tt k1}'s input has been removed. When there exists at least one completed enqueue whose value has not been removed, the rule {\tt lin(deq,y,k)} sets {\tt y} to the value added by a minimal enqueue (w.r.t. the current happens-before).

@@

\begin{minipage}[t]{7.5cm}
\vspace{-3mm}
\begin{lstlisting}
rule inv(enq,x,k) = {
  $\mathsf{present}$(k) = true;
  $\mathsf{pending}$(k) = true;
  $\mathsf{arg}$(k) = x;
  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
    $\mathsf{before}$(k1,k) = true;
}

rule lin(deq,y,k) = {
  choose a $\in$ {true, false};
  if ( $($a$\ \vee\ \forall k'.\ \mathsf{present}(k')$=false$)\ \land\ \forall k'.\ \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$ ) 
    y = EMPTY;
  else 
    choose k1 $\in$ $\mathsf{Operations}$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
    $\mathsf{present}$(k1) = false;
    y = $\mathsf{arg}$(k1);
}
\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-3mm}
\begin{lstlisting}
rule ret(enq,k) = {
  $\mathsf{pending}$(k) = false;
  return;
}
rule inv(deq,k) = {
  skip;
}
rule ret(deq,y,k) = {
  return y;
}
\end{lstlisting}
\end{minipage}
%	\caption{The abstract queue implementation $AbsQ$.}
%	\label{fig:signatures}
%\end{figure}

%{\scriptsize
%  \centering
%  \begin{mathpar}
%    \inferrule[call-enq]{
%      k\not\in dom(cp) \\ 
%      d\neq {\tt EMPTY}
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{inv(enq,d,k)}
%      %O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto 1]
%      O\cup\{k\},<\cup\ {\tt COMP}(O)\times\{k\},\ell[k\mapsto (d,{\tt PEND})],rv,cp[k\mapsto A_1]
%    }\hspace{5mm}
%
%    \inferrule[call-deq]{
%      k\not\in dom(cp) \\ 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{inv(deq,k)}
%      O,<,\ell,rv,cp[k\mapsto R_1]
%    }\hspace{5mm}
%    \inferrule[ret-deq]{
%       cp(k) = R_2 \\
%       rv(k)=d  
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(deq,d,k)}
%      O,<,\ell,rv,cp[k\mapsto R_3]
%    }\hspace{5mm}
%
%    \inferrule[ret-enq1]{
%      cp(k) = A_1 \\
%      k \in O \\
%      \ell(k) = (d,{\tt PEND}) 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(enq,k)}
%      O,<,\ell[k\mapsto (d,{\tt COMP})],rv,cp[k\mapsto A_2]
%    }\hspace{5mm}
%    \inferrule[ret-enq2]{
%      cp(k) = A_1 \\
%      k \not\in O 
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{ret(enq,k)}
%      O,<,\ell,rv,cp[k\mapsto A_2]
%    }\hspace{5mm}
%
%    \inferrule[lin-deq1]{
%       cp(k) = R_1 \\
%       d\neq{\tt EMPTY} \\
%       k'\in min(O) \\ 
%       \ell_1(k')=d
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{lin(deq,d,k)}
%      O\setminus \{k'\},<\uparrow k',\ell,rv[k\mapsto d],cp[k\mapsto R_2]
%    }\hspace{5mm}
%    \inferrule[lin-deq2]{
%       cp(k) = R_1 \\
%       \forall o\in O.\ \ell_2(o)={\tt PEND}
%    }{
%      O,<,\ell,rv,cp
%      \xrightarrow{lin(deq,{\tt EMPTY},k)}
%      O,<,\ell,rv[k\mapsto {\tt EMPTY}],cp[k\mapsto R_2]
%    }\hspace{5mm}    
%      \end{mathpar}
%  }
 \vspace{-4mm}
  \caption{The abstract queue implementations $AbsQ$. Each macro rule is executed atomically in one step (indicated by the keyword {\ttfamily \bfseries atomic}).
%  transition relation of $AbsQ$. We use the following notations: $\ell_i(k)$ denotes the projection of $\ell(k)$ over the $i$-th component, for each $i\in\{1,2\}$, ${\tt COMP}(O)=\{k\in O: \ell_2(k)={\tt COMP}\}$, $\mathit{f}[x\mapsto y]$ is the function $g$ such that $g(z)=f(z)$ for all $z\neq x$ in the domain of $f$, and $g(x)=y$, $min(O)$ is the set of elements of $O$ which are minimal in the order relation $<$, and $<\uparrow k$ denotes the relation $<$ where all the pairs containing $k$ have been removed.
  %\textcolor{red}{Call-Enq must have $d!= \texttt{EMPTY}$ as a premise. Also lin deq returning empty must be changed as before.}
  }
  \label{fig:transitions:AbsQ}
\vspace{-6mm}
\end{figure}

### 2017-05-04 draftCAV/refProofs.tex
stack implementation

The notion of $\Gamma$-refinement instantiated to libraries (i.e., to LTSs defining libraries) implies the notion of refinement in Definition~\ref{def:lib_ref}, for every $\Gamma$ such that $C\cup R \subseteq \Gamma$.
%, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions s.t. $C\cup R\subseteq $
%
%When $A_1$ and $A_2$ are libraries, $\Gamma$-refinement implies the notion of library refinement  for any $\Gamma$ as in Definition~\ref{def:gref}.

@@

%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement. 
\vspace{-1.5mm}
\begin{definition}
%Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. 
A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $A_1$ to $A_2$ if{f} $\mathit{fs}[s_0^1] = \{s_0^2 \}$ and:
\vspace{-1.5mm}
\begin{itemize}
\item If $(s,\gamma,s') \in \delta_1$, for some $\gamma\in \Gamma$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for each $j\neq i$.
\item If $(s,e,s') \in \delta_1$, for some $e \in \Sigma_1\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $\sigma\in (\Sigma_2\setminus\Gamma)^*$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\vspace{-3.5mm}
\end{definition}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation requires that every step of $A_1$ corresponds to a sequence of steps of $A_2$. To imply $\Gamma$-refinement, every step of $A_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $A_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
The dual notion of \emph{backward} simulation where steps are simulated backwards can be defined in a similar way.

@@

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. Since the abstract implementation is usually $(C\cup R\cup Lin)$-deterministic,
by Theorem~\ref{th:forSim}, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
from $L_1$ to $L_2$.

>>>>

The notion of $\Gamma$-refinement instantiated to libraries (i.e., to LTSs defining libraries) implies the notion of refinement in Definition~\ref{def:lib_ref} for every $\Gamma$ such that $C\cup R \subseteq \Gamma$.
%, resp., such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$. Also, let $\Gamma$ be a set of actions s.t. $C\cup R\subseteq $
%
%When $A_1$ and $A_2$ are libraries, $\Gamma$-refinement implies the notion of library refinement  for any $\Gamma$ as in Definition~\ref{def:gref}.

@@

%We can extend the relations between libraries by introducing simulation relations. We will later show that simulation relations imply refinement. 
\vspace{-1.5mm}
\begin{definition}
%Let $L_1=(Q_1,\Sigma, s_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, s_0^2, \delta_2)$ be two libraries over alphabets $\Sigma_1$ and $\Sigma_2$, respectively, such that $C\cup R \subseteq \Sigma_1\cap\Sigma_2$, and $\Gamma$ a set of actions such that $C\cup R\subseteq \Gamma\subseteq \Sigma_1\cap\Sigma_2$. 
A relation $F \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $A_1$ to $A_2$ if{f} $F(s_0^1,s_0^2)$ and:
\vspace{-1.5mm}
\begin{itemize}
\item For all $s,s'\in Q_1$, $\gamma\in \Gamma$, and $u\in Q_2$, such that $(s,\gamma,s') \in \delta_1$ and $F(s,u)$, we have that there exists $u'\in Q_2$ such that $F(s',u')$ and $u \xrightarrow{@s} u'$ where $@s_i=\gamma$, for some $i$, and $@s_j\in \Sigma_2\setminus\Gamma$, for all $j\neq i$.
\item For all $s,s'\in Q_1$, $e \in \Sigma_1\setminus \Gamma$, and $u\in Q_2$, such that $(s,e,s') \in \delta_1$ and $F(s,u)$, we have that there exists $u'\in Q_2$ such that $F(s',u')$ and $u \xrightarrow{\sigma} u'$ where $\sigma\in (\Sigma_2\setminus\Gamma)^*$.  
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\vspace{-3.5mm}
\end{definition}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation requires that every step of $A_1$ corresponds to a sequence of steps of $A_2$. To imply $\Gamma$-refinement, every step of $A_1$ labeled by an observable action $\gamma\in \Gamma$ should be simulated by a sequence of steps of $A_2$ where exactly one transition is labeled by $\gamma$ and all the other transitions are labeled by non-observable actions.
The dual notion of \emph{backward} simulation where steps are simulated backwards can be defined in a similar way.

@@

Proving the correctness of an implementation $L_1$ of a concurrent object such as a queue or a stack with fixed linearization points
reduces to proving that $L_1$ is a $(C\cup R\cup Lin)$-refinement of an abstract implementation $L_2$ of the same object where method
bodies are guarded with a global-lock acquisition. As a direct consequence of Theorem~\ref{th:forSim}, since the abstract implementation is 
 $(C\cup R\cup Lin)$-deterministic, proving $(C\cup R\cup Lin)$-refinement is equivalent to finding a $(C\cup R\cup Lin)$-forward simulation 
from $L_1$ to $L_2$.
